---
title: "Logistic Modeling and Plotting Curves"
subtitle:
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

### Are Emily and Greg More Employable Than Lakisha and Jamal? ###

Bertrand, M., & Mullainathan, S. (2004). Are Emily and Greg more employable than Lakisha and Jamal? A field experiment on labor market discrimination. _American Economic Review, 94_(4), pp. 991-1013.

\[Abstract\]
We perform a field experiment to measure racial discrimination in the labor market. We respond with fictitious resumes to help-wanted ads in Boston and Chicago newspapers. To manipulate perception of race, each resume is randomly assigned either a very African American sounding name or a very White sounding name. The results show significant discrimination against African-American names: White names receive 50 percent more callbacks for interviews. We also find that race affects the benefits of a better resume. For White names, a higher quality resume elicits 30 percent more callbacks whereas for African Americans, it elicits a far smaller increase. Applicants living in better neighborhoods receive more callbacks but, interestingly, this effect does not differ by race. The amount of discrimination is uniform across occupations and industries. Federal contractors and employers who list “Equal Opportunity Employer” in their ad discriminate as much as other employers. We find little evidence that our results are driven by employers inferring something other than race, such as social class, from the names. These results suggest that racial discrimination is still a prominent feature of the labor market.    


Variables     | Descriptions
-----------   | -----------------------------------------------------
_call_        | Was the applicant called back? (1 = yes; 0 = no)
_ethnicity_   | indicating ethnicity (i.e., "Caucasian-sounding" vs. "African-American sounding" first name)
_sex_         | indicating sex
_quality_     | Indicating quality of resume.
_experience_  | Number of years of work experience on the resume
_equal_       | Is the employer EOE (equal opportunity employment)?


```{r}
library(readr)
Resume <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/ResumeNames455.csv")
head(Resume)
library(Stat2Data)
```


1) Construct a logistic model to predict if the job applicant was called back using _experience_ as the predictor variable.

```{r}
mod1 = glm(call~experience, family=binomial, data=Resume)
summary(mod1)
```


2) Plot the raw data and the logistic curve on the same axes.

```{r}
plot(jitter(call,amount=0.1)~experience,data=Resume)

B0 = summary(mod1)$coef[1]
B1 = summary(mod1)$coef[2]

logit = function(B0, B1, x)
  {
    exp(B0+B1*x)/(1+exp(B0+B1*x))
}

curve(logit(B0, B1, x),add=TRUE, col="red")
```


3) For an applicant with 6 years of experience, what does your model predict is the probability of this applicant getting called back?

```{r}
newx=data.frame(experience=6)
head(newx)

predict(mod1, newx, type="response")
```
The probability of a call back with 6 years of experience is 7.41%.


4) Construct an empirical logit plot and comment on the linearity of the data.

```{r}
emplogitplot1(call~experience, data=Resume, ngroups = 9)
```

The linearity is fairly good. There are two points almost directly on the line and the rest of the points are fairly close to this line. This is overall a good fit.


5) Use the model from question #1 to perform a hypothesis test to determine if there is significant evidence of a relationship between _call_ and _experience_. Cite your hypotheses, p-value, and conclusion in context.

```{r}
summary(mod1)
```

I am using the summary test for the slope coefficient to determine if there is significant evidence of a relationship between call and experience. The null hypothesis is that the slope coefficient to experience is 0, and the alternative hypothesis is that the slope coefficient to experience is nonzero. The p-value is 2.07e-05, which is significant enough for us to conclude that the slope coefficient to experience is nonzero.

6) Construct a confidence interval for the odds ratio for your model and include a sentence interpreting the interval in the context.

```{r}
exp(confint(mod1))
```

We are 95% confident that the change in odds ratio (similar to slope) is between 1.02 and 1.06.

7) For each 2-year increase in _experience_, how does your model predict the odds will change for the applicant getting called back?

```{r}
summary(mod1)

exp(0.03908)
```

Every 2-years, my model predicts the odds will change by 2.08.

8) Construct subsets of the data for each category of _ethnicity_ and construct logistic models to predict if the job applicant was called back using _experience_ as the predictor variable for each of these subsets. Then plot the raw data and the logistic curves on the same axes. Comment on differences between the curves and what this means in the context of the data.

```{r}
resume_cauc=subset(Resume, ethnicity=='cauc')
resume_afam=subset(Resume, ethnicity=='afam')

mod_cauc = glm(call~experience, family=binomial, data=resume_cauc)
mod_afam = glm(call~experience, family=binomial, data=resume_afam)

plot(jitter(call,amount=0.1)~experience,data=Resume)

B0_cauc = summary(mod_cauc)$coef[1]
B1_cauc = summary(mod_cauc)$coef[2]

B0_afam = summary(mod_afam)$coef[1]
B1_afam = summary(mod_afam)$coef[2]

curve(logit(B0_cauc, B1_cauc, x),add=TRUE, col="red")
curve(logit(B0_afam, B1_afam, x),add=TRUE, col="blue")
```

The curve that represents the Caucasian population has a higher y-intercept and remains at a higher callback rate throughout the whole curve. The curve that represents the African American population sits below the callback curve for Caucasians. This means that the callback rate for the Caucasian population is higher, at the same levels of experience, than the African American population. 

9) Construct subsets of the data for each category of _sex_ and construct logistic models to predict if the job applicant was called back using _experience_ as the predictor variable for each of these subsets. Then plot the raw data and the logistic curves on the same axes. Comment on differences between the curves and what this means in the context of the data.

```{r}
resume_female=subset(Resume, sex=='female')
resume_male=subset(Resume, sex=='male')

mod_female = glm(call~experience, family=binomial, data=resume_female)
mod_male = glm(call~experience, family=binomial, data=resume_male)

plot(jitter(call,amount=0.1)~experience,data=Resume)

B0_female = summary(mod_female)$coef[1]
B1_female = summary(mod_female)$coef[2]

B0_male = summary(mod_male)$coef[1]
B1_male = summary(mod_male)$coef[2]

curve(logit(B0_female, B1_female, x),add=TRUE, col="green")
curve(logit(B0_male, B1_male, x),add=TRUE, col="orange")

summary(mod_male)
summary(mod_female)
```
The curve for male applicants is nearly flat, while the female curve has a much larger slope. This means that men get called back at the same rate, no matter the experience. This is not the case for females.